Disclosing funding sources for open access publication fees: the Open APC

### 2016-06-02 article.Rmd
minor fixes comparision

<!-- This complex situation of fee-based open access publishing creates difficulties for researchers and practitioners alike. Some European initiatives have therefore begun to collect and disclose their expenditures on open access journal articles as open data. To our knowledge, the first research funders providing such data were the Wellcome Trust [@wellcome_apc] and the Austrian Science Fund FWF [@fwf_apc]. The not-for-profit company Jisc followed by collecting data from higher-education institutions in the UK [@Lawson_data_2015]. Disclosed as publicly available spreadsheets, these data-sets self-report payments made along with bibliographic information, including title, journal and publisher, and a persistent identifier to the publisher's version, and a link to a deposit in a subject repository. Curatorial efforts focused on the disambiguation of publisher and journal titles as well as on detecting duplicates. In the case of the Wellcome Trust, crowd-sourcing data cleaning activities through a Google spreadsheet in combination with checks against bibliographic sources massively improved the spending data (see comments in @wellcome_apc).-->

@@

Data collection in this study also involved obtaining cost data from related open data initiatives. To compare self-reported spending on open access journal articles by Germany universities and research organisations with that of other initiatives, we reviewed the openly available data-sets from the Wellcome Trust [@wellcome_apc], the Austrian Science Fund FWF [@fwf_apc] and Jisc. For analysis, we obtained the amount of overall spending on publications fees, as well as on charges to publish in fully open access journal articles. In the case of FWF, we gathered the cost information from accompanying reports. We used the spreadsheet data to summarize  Wellcome Trust's and Jisc's spending, and converted the prices from GBP to Euro in accordance with the average Euro foreign exchange reference rates provided by the European Central Bank. Our comparison between the open data initiatives focussed on the last two years 2014 and 2015. Because Wellcome Trust's spending was reported for the periods 2013 - 2014 and 2014 - 2015, we referred to the average exchange rates of the full two-year period as we could not determine the actual invoicing dates from the data. We excluded Jisc's spending in 2014 because full currency information in the data-set was not available.

>>>>

<!-- This complex situation of fee-based open access publishing creates difficulties for researchers and practitioners alike. Some European initiatives have therefore begun to collect and disclose their expenditures on open access journal articles as open data. To our knowledge, the first research funders providing such data were the Wellcome Trust [@fwf_apc_13] and the Austrian Science Fund FWF [@wellcome_apc_12]. The not-for-profit company Jisc followed by collecting data from higher-education institutions in the UK [@Lawson_data_2015]. Disclosed as publicly available spreadsheets, these data-sets self-report payments made along with bibliographic information, including title, journal and publisher, and a persistent identifier to the publisher's version, and a link to a deposit in a subject repository. Curatorial efforts focused on the disambiguation of publisher and journal titles as well as on detecting duplicates. In the case of the Wellcome Trust, crowd-sourcing data cleaning activities through a Google spreadsheet in combination with checks against bibliographic sources massively improved the spending data (see comments in @wellcome_apc).-->

@@

Data collection also involved obtaining cost data from related open data initiatives. To compare self-reported spending on open access journal articles by Germany universities and research organisations with that of other initiatives, we reviewed the openly available data-sets from the the Austrian Science Fund FWF [@fwf_apc_14; @fwf_apc_15], Jisc [@jisc_14; @jisc_15] and the Wellcome Trust [@wellcome_apc_13_14; @wellcome_apc_14_15]. For analysis, we obtained the overall publication fee spending to support fully and hybrid open access journal articles. In the case of FWF, we gathered the cost information from the accompanying spending reports. We used the spreadsheet data to calculate  Wellcome Trust's and Jisc's spending, and converted the prices from GBP to Euro in accordance with the average Euro foreign exchange reference rates provided by the European Central Bank. Our comparison between the open data initiatives focussed on the last two years 2014 and 2015. Because Wellcome Trust's spending was reported for the periods 2013 - 2014 and 2014 - 2015, we referred to the average exchange rates of the full two-year period as we could not determine the actual invoicing dates from the data. We excluded articles from the analysis for which neither price information nor the journal type could be obtained. In the case of Jisc's 2014 data [@jisc_14], 4.861.772 € being spent on 2,812 publications were not eligible for further analysis because no publication type was given in the data-set.
